The Numerical Accuracy of Computing Electrostatic Interaction in the Inhomogeneous and Correlated Molecular Systems: for the Ewald Summation, SPME and Staggered Mesh Ewald Methods.

### 2012-07-03 longRange.hetero/alltext/publish/resubmit/lett.tex
fix typos, add fig.rho.eps

Reviewer  1 Comments:

At a glance, the paper by Han Wang and colleagues appears to be another error estimator for P3M and related electrostatics methods.  But, after taking in the second paragraph it becomes clear that the authors are much more than mathematically inclined on the subject, and have indeed performed a useful service to the simulations community by incorporating the correlation of molecular charges into their error estimates.  It is probably not of tremendous practical benefit in an application like an Ewald parameter optimizer, as these programs can just repeat an electrostatic calculation on a representative snapshot of a simulation and estimate the error directly from the resulting forces.  However, it is of value to have a formal expression for the effect of charge correlations on these errors, and it may be instrumental in devising a pairwise approximate correction to electrostatic forces obtained from coarse grids.  The paper will require minor modifications so that the authors do not run afoul of various members of the community who hold strong opinions, but otherwise appears to be fit for publication.

In the introduction, the authors should probably take the perspective that SPME is a special case of P3M.  In the opinions of the developers of SPME, the derivation differs from P3M although it leads to similar results, but even TA Darden acknowledges that the reciprocal space component of the non-smooth PME algorithm was inspired by P3M.  It is probably best to phrase SPME as such and not offend anyone.  The interlacing technique was first applied in the 1970s, as the authors have cited.  It was originally a means of reducing the memory requirements of these calculations—electrostatics on the two grids were solved serially.  Finally, the sentence “We prove that the error is decomposed…” should be changed to “We show that the error may be approximately decomposed…” to indicate that the estimator is, in the end, approximate but relevant to real systems.

Cosmetically, the authors should combine the two figures showing the charge distributions of Examples 1 and 2, to ensure they appear side-by-side in the published version.

@@

page 3 lines 9,14  The term "research" could be replaced by "papers" on line 9 and "paper" on 14
another possibility is "manuscript"

page 3 line 27 "which assume the error" $\rightarrow$ "which assumes the error"
page 3 line 43 "nearest neighbor approximation of correlation error": insert "the" before "correlation"
page 3 line 54 "we consider the system" $\rightarrow$ "we consider a system"
page 3 line 55 "is composed by N" $\rightarrow$ "is composed of N"
page 3 line 56 "the force of a" $\rightarrow$ "the force on a"
page 4 line 18 "error estimate for the error force" seems like 1st "error' is unnecessary
              perhaps replace with "following estimate for the error force"---also the term "error force" could be
              replaced with "force error" here and elsewhere and I think the text would flow better
page 6 line 16 "Most of the error estimates" $\rightarrow$ "Most previous error estimates"
page 6 line 32: "Analog to" $\rightarrow$ "Analogous to"
page 7 line 3 "A non-varnishing" $\rightarrow$ "A non-vanishing"
page 7 line 10,11: "the locally neutral condition" $\rightarrow$ "the local neutrality condition"
page 7 line 21: "which is still not applicable" $\rightarrow$ "which is still not tractable" or maybe "computationally feasible"
page 7 line 46: "Then the second part" $\rightarrow$ "Then the second term". Also on line 54 "part" $\rightarrow$ "term"
page 7 line 52: "small comparing with" $\rightarrow$ "small compared with" or "small in comparison with". same for pg 8 line 16
page 8 line 42: "Ensemble averages is taken" $\rightarrow$ "Ensemble averages are taken"
page 10 line 1: "ia the reciprocal" $\rightarrow$ "are the reciprocal"
page 11 line 8: "becomes not applicable" as in page 7 line 21 I think "intractable" is better here
page 11 line 10: "PME ans SPME" ; "ans" $\rightarrow$ "and"
page 12 line 7: "only little" $\rightarrow$ "only a small"
page 14 line 33: "It is shown that the self-interaction error dominate" $\rightarrow$ "It has been shown that the sel-interaction
                   error dominates"
page 17 line 49: "Twice as larger" $\rightarrow$ "Twice as large"
page 18 line 43: "cut-offed" $\rightarrow$ "truncated"
page 18 line 47: "are consistent very well with" $\rightarrow$ "are consistently accurate compared with" or "accurately track"
page 18 line 60: "reduces by more than" $\rightarrow$ "is reduced by more than"
page 19 line 34: "FFTs than" $\rightarrow$ "FFTs compared to"
page 19 line 55: again "cut-offed" $\rightarrow$ "truncated"
page 20 line 34: "was reported" $\rightarrow$ "were reported"
page 20 line 35: "varnish" $\rightarrow$ "vanish"
page 20 line 41: "error estimates consist with" $\rightarrow$ "error estimates are consistent with" or
                                                  "error estimates accurately track"

page 23 line 57: "composed by three additive" $\rightarrow$ "composed of three additive"
page 24 line 11: "the error is of good property" Not clear what is meant. Do you mean total error is not
                 contaminated by this?
page 24 line 18: "was proved to dominant" $\rightarrow$ "was proved to dominate"
page 24 line 19: "helped removing" $\rightarrow$ "helped to remove"
page 24 line 35: "only one half FFTs" $\rightarrow$ "only one half the FFTs"
page 27, line 15: "Inserting it into Eqn (17), also neglect" $\rightarrow$ "Inserting this into Eqn (17) while neglecting"

>>>>

\noindent
Reviewer  1 Comments:\\

\textit{ At a glance, the paper by Han Wang and colleagues appears to
  be another error estimator for P3M and related electrostatics
  methods.  But, after taking in the second paragraph it becomes clear
  that the authors are much more than mathematically inclined on the
  subject, and have indeed performed a useful service to the
  simulations community by incorporating the correlation of molecular
  charges into their error estimates.  It is probably not of
  tremendous practical benefit in an application like an Ewald
  parameter optimizer, as these programs can just repeat an
  electrostatic calculation on a representative snapshot of a
  simulation and estimate the error directly from the resulting
  forces.  However, it is of value to have a formal expression for the
  effect of charge correlations on these errors, and it may be
  instrumental in devising a pairwise approximate correction to
  electrostatic forces obtained from coarse grids.  The paper will
  require minor modifications so that the authors do not run afoul of
  various members of the community who hold strong opinions, but
  otherwise appears to be fit for publication.
}

?

\textit{
In the introduction, the authors should probably take the perspective that SPME is a special case of P3M.  In the opinions of the developers of SPME, the derivation differs from P3M although it leads to similar results, but even TA Darden acknowledges that the reciprocal space component of the non-smooth PME algorithm was inspired by P3M.  It is probably best to phrase SPME as such and not offend anyone.  The interlacing technique was first applied in the 1970s, as the authors have cited.  It was originally a means of reducing the memory requirements of these calculations—electrostatics on the two grids were solved serially.  Finally, the sentence “We prove that the error is decomposed…” should be changed to “We show that the error may be approximately decomposed…” to indicate that the estimator is, in the end, approximate but relevant to real systems.
}\\

not done\\

\textit{
Cosmetically, the authors should combine the two figures showing the charge distributions of Examples 1 and 2, to ensure they appear side-by-side in the published version.
}\\

@@

page 3 lines 9,14  The term "research" could be replaced by "papers" on line 9 and "paper" on 14
another possibility is "manuscript"
page 3 line 27 "which assume the error" $\rightarrow$ "which assumes the error"
page 3 line 43 "nearest neighbor approximation of correlation error": insert "the" before "correlation"
page 3 line 54 "we consider the system" $\rightarrow$ "we consider a system"
page 3 line 55 "is composed by N" $\rightarrow$ "is composed of N"
page 3 line 56 "the force of a" $\rightarrow$ "the force on a"
*page 4 line 18 "error estimate for the error force" seems like 1st "error' is unnecessary
              perhaps replace with "following estimate for the error force"---also the term "error force" could be
              replaced with "force error" here and elsewhere and I think the text would flow better
page 6 line 16 "Most of the error estimates" $\rightarrow$ "Most previous error estimates"
page 6 line 32: "Analog to" $\rightarrow$ "Analogous to"
page 7 line 3 "A non-varnishing" $\rightarrow$ "A non-vanishing"
page 7 line 10,11: "the locally neutral condition" $\rightarrow$ "the local neutrality condition"
page 7 line 21: "which is still not applicable" $\rightarrow$ "which is still not tractable" or maybe "computationally feasible"
page 7 line 46: "Then the second part" $\rightarrow$ "Then the second term". Also on line 54 "part" $\rightarrow$ "term"
page 7 line 52: "small comparing with" $\rightarrow$ "small compared with" or "small in comparison with". same for pg 8 line 16
page 8 line 42: "Ensemble averages is taken" $\rightarrow$ "Ensemble averages are taken"
page 10 line 1: "ia the reciprocal" $\rightarrow$ "are the reciprocal"
page 11 line 8: "becomes not applicable" as in page 7 line 21 I think "intractable" is better here
page 11 line 10: "PME ans SPME" ; "ans" $\rightarrow$ "and"
page 12 line 7: "only little" $\rightarrow$ "only a small"
page 14 line 33: "It is shown that the self-interaction error dominate" $\rightarrow$ "It has been shown that the sel-interaction
                   error dominates"
page 17 line 49: "Twice as larger" $\rightarrow$ "Twice as large"
page 18 line 43: "cut-offed" $\rightarrow$ "truncated"
page 18 line 47: "are consistent very well with" $\rightarrow$ "are consistently accurate compared with" or "accurately track"
page 18 line 60: "reduces by more than" $\rightarrow$ "is reduced by more than"
page 19 line 34: "FFTs than" $\rightarrow$ "FFTs compared to"
page 19 line 55: again "cut-offed" $\rightarrow$ "truncated"
page 20 line 34: "was reported" $\rightarrow$ "were reported"
page 20 line 35: "varnish" $\rightarrow$ "vanish"
page 20 line 41: "error estimates consist with" $\rightarrow$ "error estimates are consistent with" or
                                                  "error estimates accurately track"

page 23 line 57: "composed by three additive" $\rightarrow$ "composed of three additive"
*page 24 line 11: "the error is of good property" Not clear what is meant. Do you mean total error is not
                 contaminated by this?
page 24 line 18: "was proved to dominant" $\rightarrow$ "was proved to dominate"
page 24 line 19: "helped removing" $\rightarrow$ "helped to remove"
page 24 line 35: "only one half FFTs" $\rightarrow$ "only one half the FFTs"
page 27, line 15: "Inserting it into Eqn (17), also neglect" $\rightarrow$ "Inserting this into Eqn (17) while neglecting"
